---
title: "Behaviorals for Synapse"
author: "Lindsey Hasak"
date: "2/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# load libraries
library(Hmisc)      # for correlation matrix
library(psych)      # for descriptive statistics
library(knitr)      # for knitting to pdf
library(kableExtra) # for making pretty tables
library(ggpubr)     # for arrangling plots
library(corrplot)   # for making correlogram
library(forcats)    # for reordering factor levels
library(tidyverse)  # for data wrangling
```

## Load and recode; summary statistics
```{r data cleaning, echo = FALSE, message = FALSE}
# load the data and indicate how NAs are entered
b_data <- read_csv("K2Behavioral_T1.csv", col_names = TRUE, na = c("N/A"))
colnames(b_data)

b_clean <- b_data %>% 
  # rename columns for consistency 
  rename(part_id = participant_id,
         ctopp_c_raw = ctopp_color_raw,
         ctopp_c_scale = ctopp_color_scaled,
         ctopp_l_raw = ctopp_lttr_raw,
         ctopp_l_scale = ctopp_lttr_scaled,
         swe_raw = swe_raw2,
         pde_raw = pde_raw_final,
         swe_age = twre_scaled_age,
         pde_age = pde_scaled_age,
         swe_grade = swe_scaled_grade,
         pde_grade = pde_scaled_grade,
         wj_raw = wj_corr_final,
         wj_age_b = wj_ageeq_basal,
         wj_grade_b = wj_gradeeq_basal) %>% 
  # separate age column into year, month, and day columns
  separate(col = k2_age,
           into = c("year", "month", "day"),
           sep = "-") %>% # indicate separating character
  # separate grade column into grade and months in grade
  separate(col = k2_grade,
           into = c("grade", "g_month"),
           sep = "_") %>% 
  # reorders grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,
                             "K", "1", "2")) %>% 
  # convert most columns to doubles (numeric values)
  type_convert()
  
# now age back as new column in decimals
b_clean$age <- b_clean$year + (b_clean$month/12) + (b_clean$day/365)

# and finally look at data structure
head(b_clean)

# still need to edit this to make sense of WJ age and grade equivalents
```


```{r describe stats, message = FALSE, echo = FALSE, warning = FALSE}
# deselect unecessary columns and get summary statistics
b_clean_descr <- b_clean %>% 
  select(- part_id,
         - k2behav_examiner,
         - k2_hand,
         - k2behav_notes,
         - year,
         - month,
         - day,
         - grade,
         - g_month,
         # leave out these two columns for now
         - twre_scaledsc_age,
         - twre_scaledsc_grade) %>% 
  # puts age first
  select(age, everything())

# put output in a nice table
b_clean_descr <- data.frame(describe(b_clean_descr))

# kable is a simple table generator
kable(b_clean_descr, 
      digits = 3, # 3 digits (or however many you want)
      format = "latex", 
      caption = "K2 Behavioral Descriptive Statistics") %>% 
  kable_styling(latex_options = c("striped", "scale_down")) # from the kableExtra package
```

# CTOPP 

### Density plot - all grades

Current colors for the mean lines are wrong and need to be fixed.

```{r ctopp c density plot}

ggplot(b_clean, aes(ctopp_c_raw)) +
  geom_density(aes(fill = grade), alpha = 0.3) +
  geom_vline(data = b_clean %>% 
               filter(grade == "K"), 
             aes(xintercept = mean(na.omit(ctopp_c_raw))),
             color = "red", 
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "1"), 
             aes(xintercept = mean(na.omit(ctopp_c_raw))),
             color = "darkgreen",
             alpha = 0.5,
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "2"), 
             aes(xintercept = mean(na.omit(ctopp_c_raw))),
             color = "blue",
             linetype="dashed",
             size = 0.5,
             alpha = 0.6) +
  labs(x = "color naming (seconds)",
       y = "probability of a student obtaining this score") +
  scale_x_continuous(limits = c(0, 90),
                     breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80)) +
  scale_y_continuous(expand = expand_scale(mult = c(0, .1))) +
  theme_classic()

ggsave("ctopp_density_all.pdf")
```

```{r ctopp l density plot }
ggplot(b_clean, aes(ctopp_l_raw)) +
  geom_density(aes(fill = grade), alpha = 0.3) +
  geom_vline(data = b_clean %>% 
               filter(grade == "K"), 
             aes(xintercept = mean(na.omit(ctopp_l_raw))), 
             color = "red", 
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "1"), 
             aes(xintercept = mean(na.omit(ctopp_l_raw))), 
                 color = "darkgreen",
             alpha = 0.5,
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "2"), 
             aes(xintercept = mean(na.omit(ctopp_l_raw))), 
             color = "blue",
             linetype="dashed",
             size = 0.5,
             alpha = 0.6) +
  scale_x_continuous(limits = c(0, 90),
                     breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80)) +
  scale_y_continuous(expand = expand_scale(mult = c(0, .1))) +
  labs(x = "letter naming (seconds)",
       y = "probability of a student obtaining this score") +
  xlim(0, 85) +
  theme_classic()

ggsave("ctoppl_density_all.pdf")
```


### CTOPP Density Plots by Grade - Kindergarten

Lines in the first plot denote letter naming norms for 6-year olds, *M* = 35, *SD* = 17. Lines in the second plot denote color naming norms for 6-year olds, *M* = 41, *SD* = 16. Mean age of Kindergarten class at time of testing was 5.92 years. 

```{r}
l1 <- b_clean %>% 
  filter(grade == "K") %>%
  ggplot(aes(ctopp_l_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 35),
            color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 18),
            color="black", linetype="dashed", size = 0.5) +
    geom_vline(aes(xintercept = 52),
            color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
   scale_y_continuous(limits = c(0, 14),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Letter Fluency - Kindergarten",
       x = "Letter Naming (s)",
       y = "# of Students") +
  theme_classic()

ggsave("ctoppl_kindergarten.pdf")
```

```{r ctopp color k}

c1 <- b_clean %>% 
  filter(grade == "K") %>%
  ggplot(aes(ctopp_c_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 41),
             color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 25),
             color="black", linetype="dashed", size = 0.5) +
  geom_vline(aes(xintercept = 57),
             color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
  scale_y_continuous(limits = c(0, 14),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Color Fluency - Kindergarten", 
       x = "Color Naming (s)",
       y = "# of Students") +
  theme_classic()

ggsave("ctoppc_kindergarten.pdf")
```

### First grade
Lines in the first plot denote letter naming norms for 7-year olds, *M* = 23, *SD* = 8. Lines in the second plot denote color naming norms for 6-year olds, *M* = 41, *SD* = 16 (there are no color naming norms past age 6). Mean age of the first grade class at time of testing was 6.78 years. 

```{r ctopp l first grade}
# used 7 year old norms - could also use 6 for consistency
l2 <- b_clean %>% 
  filter(grade == "1") %>%
  ggplot(aes(ctopp_l_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 23),
             color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 15),
             color="black", linetype="dashed", size = 0.5) +
  geom_vline(aes(xintercept = 31),
             color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
   scale_y_continuous(limits = c(0, 14),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Letter Fluency - First Grade",
       x = "Letter Naming (s)",
       y = "# of Students") +
  theme_classic()

ggsave("ctoppl_first.pdf")
```

```{r ctopp c first grade}
# no norms for color naming past 6 yrs old so used 6 yr old norms bc many students are still 6
c2 <- b_clean %>% 
  filter(grade == "1") %>%
  ggplot(aes(ctopp_c_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 41),
             color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 25),
             color="black", linetype="dashed", size = 0.5) +
  geom_vline(aes(xintercept = 57),
             color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
   scale_y_continuous(limits = c(0, 14),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Color Fluency - First Grade",
       x = "Color Naming (s)",
       y = "# of Students") +
  theme_classic()

ggsave("ctoppc_first.pdf")
# mean age = 6.78
  # 2nd - mean age = 7.52
```

### Second grade

Lines in the first plot denote letter naming norms for 8-year olds, *M* = 20, *SD* = 5. Mean age of the second grade class at time of testing was 7.52 years. 
```{r ctopp l second}
# used 8 year old norms - could also use 7
l3 <- b_clean %>% 
  filter(grade == "2") %>% 
  ggplot(aes((ctopp_l_raw))) +
  geom_histogram(binwidth = 5, 
                 na.rm = TRUE,
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 20),
            color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 15),
            color="black", linetype="dashed", size = 0.5) +
    geom_vline(aes(xintercept = 25),
            color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
  scale_y_continuous(limits = c(0, 14),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Letter Fluency - Second Grade",
       x = "Letter Naming (s)",
       y = "# of Students") +
  theme_classic()

ggsave("ctoppl_second.pdf")
```

```{r ctopp c second}
# no norms for color naming past 6 yrs old 
c3 <- b_clean %>% 
  filter(grade == "2") %>%
  ggplot(aes(ctopp_c_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  xlim(0, 80) +
  scale_y_continuous(limits = c(0, 14),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Color Fluency - Second",
       x = "Color Naming (s)",
       y = "# of Students") +
  theme_classic()

ggsave("ctoppc_second.pdf")

```

```{r ctopp hist pdfs}
ctoppc <- ggarrange(c1, c2, c3,
                  ncol = 1,
                  nrow = 3,
                  legend = "none")
ggsave("ctoppc_hist_all.pdf", ctoppc)

ctoppl <- ggarrange(l1, l2, l3,
                  ncol = 1,
                  nrow = 3,
                  legend = "none")
ggsave("ctoppl_hist_all.pdf", ctoppl)
```


# TOWRE Histograms
Norms from chart on page 85

```{r swe kindergarten}
swe1 <- b_clean %>% 
  filter(grade == "K") %>%
  ggplot(aes(swe_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 29),
            color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 13),
            color="black", linetype="dashed", size = 0.5) +
    geom_vline(aes(xintercept = 42),
            color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
  scale_y_continuous(limits = c(0, 6),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Sight Word Fluency - Kindergarten",
       x = "",
       y = "# of Students") +
  theme_classic()

ggsave("swe_k.pdf")
```

```{r pde kindergarten}
pde1 <- b_clean %>% 
  filter(grade == "K") %>%
  ggplot(aes(pde_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 13),
             color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 4),
             color="black", linetype="dashed", size = 0.5) +
  geom_vline(aes(xintercept = 22),
             color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
 scale_y_continuous(limits = c(0, 6),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Pseudoword Fluency - Kindergarten",
       x = "",
       y = "# of Students") +
  theme_classic()

ggsave("pde_k.pdf")
```
### First grade TOWRE histograms

```{r swe first}
swe2 <- b_clean %>% 
  filter(grade == "1") %>%
  ggplot(aes(swe_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 47),
            color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 28),
            color="black", linetype="dashed", size = 0.5) +
    geom_vline(aes(xintercept = 66),
            color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
  scale_y_continuous(limits = c(0, 6),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Sightword Fluency - First Grade",
       x = "",
       y = "# of Students") +
  theme_classic()

ggsave("swe_first.pdf")
```

```{r pde first}
pde2 <- b_clean %>% 
  filter(grade == "1") %>%
  ggplot(aes(pde_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 24),
             color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 10),
             color="black", linetype="dashed", size = 0.5) +
  geom_vline(aes(xintercept = 38),
             color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
  scale_y_continuous(limits = c(0, 6),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Pseudoword Fluency - First Grade",
       x = "",
       y = "# of Students") +
  theme_classic()

ggsave("pde_first.pdf")
```

```{r swe second}
swe3 <- b_clean %>% 
  filter(grade == "2") %>%
  ggplot(aes(swe_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 55),
            color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 42),
            color="black", linetype="dashed", size = 0.5) +
    geom_vline(aes(xintercept = 68),
            color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
  scale_y_continuous(limits = c(0, 6),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Sight Word Fluency - Second Grade",
       x = "# of Sight Words Read in 45 seconds",
       y = "# of Students") +
  theme_classic()

ggsave("swe_second.pdf")
```

```{r pde second}
pde3 <- b_clean %>% 
  filter(grade == "2") %>%
  ggplot(aes(pde_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  geom_vline(aes(xintercept = 27),
             color="black", linetype="dashed", size = 1) +
  geom_vline(aes(xintercept = 16),
             color="black", linetype="dashed", size = 0.5) +
  geom_vline(aes(xintercept = 38),
             color="black", linetype="dashed", size = 0.5) +
  xlim(0, 80) +
  scale_y_continuous(limits = c(0, 6),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Pseudoword Fluency - Second Grade",
       x = "# of Pseudowords read in 45 seconds",
       y = "# of Students") +
  theme_classic()

ggsave("pde_second.pdf")
```

```{r}
swe <- ggarrange(swe1, swe2, swe3,
                  ncol = 1,
                  nrow = 3,
                  legend = "none")
ggsave("swe_hist_all.pdf", swe)

pde <- ggarrange(pde1, pde2, pde3,
                 ncol = 1,
                 nrow = 3,
                 legend = "none")
ggsave("pde_hist_all.pdf", pde)
```




```{r towre means per grade and norms}
b_clean %>% 
  filter(grade == "K") %>% 
  summarise(swe_raw = mean(na.omit(swe_raw)),
            pde_raw = mean(na.omit(pde_raw)),
            age = mean(age))
# mean swe - 9.83
# mean pde = 5.76
# mean age = 5.92
# 6 yr old swe norm - mean 29; sd = 16
# 6 yr old pde norm - m 13; sd 9

b_clean %>% 
  filter(grade == "K",
         age >= 6) %>% 
  summarise(swe_raw = mean(na.omit(swe_raw)),
            pde_raw = mean(na.omit(pde_raw)),
            age = mean(age))
# swe mean = 10.25
# pde mean = 7
# age = 6.29

b_clean %>% 
  filter(grade == "1") %>% 
   summarise(swe_raw = mean(swe_raw),
            pde_raw = mean(pde_raw),
            age = mean(age))

# mean swe - 44
# mean pde = 19.39
# mean age = 6.78
# 7 yr old swe norm - mean 47; sd = 19
# 7 yr old pde norm - m 24; sd - 14

b_clean %>% 
  filter(grade == "2") %>% 
    summarise(swe_raw = mean(swe_raw),
            pde_raw = mean(pde_raw),
            age = mean(age))

# mean swe - 55.4
# mean pde = 27.1
# mean age = 7.5
# 8 yr old swe norm - mean 55; sd = 13
# 8 yr old pde norm - m 27; sd 11

```


```{r towre swe hist, echo = FALSE, message = FALSE, warning = FALSE}
# TOWRE_SWE 
ggplot(b_clean, aes(swe_raw)) +
  geom_density(aes(fill = grade), alpha = 0.3) +
  geom_vline(data = b_clean %>% 
               filter(grade == "K"), 
             aes(xintercept = mean(na.omit(swe_raw))),
             color = "red", 
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "1"), 
             aes(xintercept = mean(na.omit(swe_raw))),
             color = "darkgreen",
             alpha = 0.7,
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "2"), 
             aes(xintercept = mean(na.omit(swe_raw))),
             color = "blue",
             linetype="dashed",
             size = 0.5,
             alpha = 0.6) +
  labs(title = "Sightword Fluency Distribution",
       x = "# Correct",
       y = "probability of a student obtaining this score") +
  xlim(0, 80) +
  scale_y_continuous(expand = expand_scale(mult = c(0, .1))) +
  theme_classic()

ggsave("swe_density_all.pdf")
```

```{r}
# TOWRE_PDE
ggplot(b_clean, aes(pde_raw)) +
  geom_density(aes(fill = grade), alpha = 0.3) +
  geom_vline(data = b_clean %>% 
               filter(grade == "K"), 
             aes(xintercept = mean(na.omit(pde_raw))),
             color = "red", 
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "1"), 
             aes(xintercept = mean(na.omit(pde_raw))),
             color = "darkgreen",
             alpha = 0.7,
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "2"), 
             aes(xintercept = mean(na.omit(pde_raw))),
             color = "blue",
             linetype="dashed",
             size = 0.5,
             alpha = 0.6) +
  labs(title = "Pseudoword Fluency Distribution",
       x = "# Correct",
       y = "probability of a student obtaining this score") +
  scale_y_continuous(expand = expand_scale(mult = c(0, .1))) +
  xlim(0, 80) +
  theme_classic()

ggsave("pde_density_all.pdf")
```

# WJ

```{r wj means per grade}
b_clean %>% 
  filter(grade == "K") %>% 
  summarise(wj_raw = mean(wj_raw),
            age = mean(age))
# m = 18.05
# 18 is K.3
# kindergarten just about what you'd expect at K.3
# would expect that slope from beginning of end to kindergarten would be higher than national average slope

b_clean %>% 
  filter(grade == "1") %>% 
  summarise(wj_raw = mean(wj_raw),
            age = mean(age))
# m = 42.17
# 42 is 2.0
# scoring near the beginning of 2nd grade

b_clean %>% 
  filter(grade == "2") %>% 
  summarise(wj_raw = mean(wj_raw),
            age = mean(age))
# m = 52.1
# 52 is 3.5
# start with WJ, show means for ages and say that they're acting like grade equivalent 3.5
```


```{r}
ggplot(b_clean, aes(wj_raw)) +
  geom_density(aes(fill = grade), alpha = 0.3) +
    geom_vline(data = b_clean %>% 
               filter(grade == "K"), 
             aes(xintercept = mean(na.omit(wj_raw))),
             color = "red", 
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "1"), 
             aes(xintercept = mean(na.omit(wj_raw))),
             color = "darkgreen",
             alpha = 0.7,
             linetype="dashed",
             size = 0.5) +
  geom_vline(data = b_clean %>% 
               filter(grade == "2"), 
             aes(xintercept = mean(na.omit(wj_raw))),
             color = "blue",
             linetype="dashed",
             size = 0.5,
             alpha = 0.6) +
  labs(title = "Letter-Word Identification Distribution",
       x = "# Correct",
       y = "probability of a student obtaining this score") +
  scale_y_continuous(expand = expand_scale(mult = c(0, .1))) +
  scale_x_continuous(limits = c(0, 70),
                     breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
  theme_classic()

ggsave("wj_density_all.pdf")
```
### Letter Word ID by Grade

```{r letter word id hist}
b_clean %>% 
  filter(grade == "K") %>%
  ggplot(aes(wj_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  xlim(0, 70) +
  scale_y_continuous(limits = c(0, 10),
                     breaks = c(2, 4, 6, 8, 10),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Letter-Word ID - K",
       x = "",
       y = "# of Students") +
  theme_classic()

b_clean %>% 
  filter(grade == "1") %>%
  ggplot(aes(wj_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  xlim(0, 70) +
  scale_y_continuous(limits = c(0, 10),
                     breaks = c(2, 4, 6, 8, 10),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Letter-Word ID - 1",
       x = "",
       y = "# of Students") +
  theme_classic()


b_clean %>% 
  filter(grade == "2") %>%
  ggplot(aes(wj_raw)) +
  geom_histogram(binwidth = 5, 
                 color = "black",
                 fill = "lightblue") +
  xlim(0, 70) +
  scale_y_continuous(limits = c(0, 10),
                     breaks = c(2, 4, 6, 8, 10),
                     expand = expand_scale(mult = c(0, .1))) +
  labs(title = "Letter-Word ID - 2",
       x = "# Correct",
       y = "# of Students") +
  theme_classic()
```


# Correlation Matrix

```{r create data for corr matrix, echo = FALSE, warning = FALSE}
# select only numeric columns to use for correlation matrix (add WJ scaled scores later)
b_corr <- b_clean %>% 
  select(ctopp_c_raw,
         ctopp_c_scale,
         ctopp_l_raw,
         ctopp_l_scale,
         swe_raw,
         swe_age,
         swe_grade,
         pde_raw,
         pde_age,
         pde_grade,
         wj_raw,
         age) %>% 
  na.omit()
```

```{r correlation matrix, echo = FALSE, warning = FALSE}

# create correlation matrix (cor from package corrplot)
b_matrix <- cor(b_corr)

# something I found online to be able to see values in correlation matrix 
library(Hmisc)
res <- rcorr(as.matrix(b_corr))

# custom function I copied and pasted - need to find link
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

```

```{r correlogram, echo = FALSE, warning = FALSE}

raw_perform <- b_corr %>% 
  select(ctopp_c_raw,
         ctopp_l_raw,
         swe_raw,
         pde_raw,
         wj_raw,
         age) %>% 
  mutate(ctopp_c_raw = ctopp_c_raw*(-1),
         ctopp_l_raw = ctopp_l_raw*(-1))

raw_matrix <- cor(raw_perform)

raw_test <- cor.mtest(raw_perform, conf.level = .95)
```

```{r}
corrplot(raw_matrix, 
         p.mat = raw_test$p, 
         insig = "label_sig",
         sig.level = c(.001, .01, .05),
         pch.cex = .7, 
         pch.col = "white",
         tl.col = "black")
```

```{r}
b_clean %>% 
  group_by(grade) %>% 
  summarise(n = n())
```

