---
title: "K2_Behavioral_T1"
author: "Lindsey Hasak"
date: "1/14/2020"
output: pdf_document
---


```{r setup, include=FALSE}
# load libraries
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)      # for correlation matrix
library(psych)      # for descriptive statistics
library(knitr)      # for knitting to pdf
library(kableExtra) # for making pretty tables
library(ggpubr)     # for arrangling plots
library(corrplot)   # for making correlogram
library(forcats)    # for reordering factor levels
library(tidyverse)  # for data wrangling
```

## Load and recode; summary statistics
```{r data cleaning, echo = FALSE, message = FALSE}
# load the data and indicate how NAs are entered
b_data <- read_csv("K2Behavioral_T1.csv", col_names = TRUE, na = c("N/A"))
colnames(b_data)

b_clean <- b_data %>% 
  # rename columns for consistency 
  rename(part_id = participant_id,
         ctopp_c_raw = ctopp_color_raw,
         ctopp_c_scale = ctopp_color_scaled,
         ctopp_l_raw = ctopp_lttr_raw,
         ctopp_l_scale = ctopp_lttr_scaled,
         swe_raw = swe_raw2,
         pde_raw = pde_raw_final,
         swe_age = twre_scaled_age,
         pde_age = pde_scaled_age,
         swe_grade = swe_scaled_grade,
         pde_grade = pde_scaled_grade,
         wj_raw = wj_corr_final,
         wj_age_b = wj_ageeq_basal,
         wj_grade_b = wj_gradeeq_basal) %>% 
  # separate age column into year, month, and day columns
  separate(col = k2_age,
           into = c("year", "month", "day"),
           sep = "-") %>% # indicate separating character
  # separate grade column into grade and months in grade
  separate(col = k2_grade,
           into = c("grade", "g_month"),
           sep = "_") %>% 
  # reorders grade variable to show in this order for plotting
  mutate(grade = fct_relevel(grade,
                             "K", "1", "2")) %>% 
  # convert most columns to doubles (numeric values)
  type_convert()
  
# now age back as new column in decimals
b_clean$age <- b_clean$year + (b_clean$month/12) + (b_clean$day/365)

# and finally look at data structure
head(b_clean)

# still need to edit this to make sense of WJ age and grade equivalents
```


```{r describe stats, message = FALSE, echo = FALSE, warning = FALSE}
# deselect unecessary columns and get summary statistics
b_clean_descr <- b_clean %>% 
  select(- part_id,
         - k2behav_examiner,
         - k2_hand,
         - k2behav_notes,
         - year,
         - month,
         - day,
         - grade,
         - g_month,
         # leave out these two columns for now
         - twre_scaledsc_age,
         - twre_scaledsc_grade) %>% 
  # puts age first
  select(age, everything())

# put output in a nice table
b_clean_descr <- data.frame(describe(b_clean_descr))

# kable is a simple table generator
kable(b_clean_descr, 
      digits = 3, # 3 digits (or however many you want)
      format = "latex", 
      caption = "K2 Behavioral Descriptive Statistics") %>% 
  kable_styling(latex_options = c("striped", "scale_down")) # from the kableExtra package
```


# Visualization

## CTOPP 

### CTOPP Histograms
```{r ctopp c histograms, echo = FALSE, message = FALSE, warning = FALSE}
#Naming convention: hp_cc = histogram plot ctopp color, etc. 
h_cc1 <- ggplot(b_clean, aes(ctopp_c_raw)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 2,
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(x = "CTOPP Color Raw Score (s)",
       y = "Density") +
  theme_classic()

h_cc2 <- ggplot(b_clean, aes(ctopp_c_scale)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 1, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(x = "CTOPP Color Raw Score (s)",
       y = "Density") +
  theme_classic()

hp_cc <- ggarrange(h_cc1, h_cc2, 
                  ncol = 2,
                  nrow = 1, 
                  common.legend = TRUE)
hp_cc

```

```{r ctopp l hist, echo = FALSE, message = FALSE, warning = FALSE}
# CTOPP L RAW
h_cl1 <- ggplot(b_clean, aes(ctopp_l_raw)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 2, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(x = "CTOPP Letter Raw Score (s)",
       y = "Density") +
  xlim(0, 85) +
  theme_classic()


# CTOPP L SCALE
h_cl2 <- ggplot(b_clean, aes(ctopp_l_scale)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 1, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(x = "CTOPP Letter Raw Score (s)",
       y = "Density") +
  theme_classic()

hp_cl <- ggarrange(h_cl1, h_cl2, 
                   ncol = 2,
                   nrow = 1, 
                   common.legend = TRUE)
hp_cl
```

### CTOPP Boxplots

```{r ctopp boxplots, echo = FALSE, message = FALSE, warning = FALSE}
b_cc1 <- ggplot(b_clean, aes(grade, ctopp_c_raw), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(title = "CTOPP Color",
       x = "Grade",
       y = "Raw Score (s)") +
  theme_classic()

b_cl1 <- ggplot(b_clean, aes(grade, ctopp_l_raw), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(title = "CTOPP Letter",
       x = "Grade",
       y = "Raw Score (s)") +
  theme_classic()

b_cc2 <- ggplot(b_clean, aes(grade, ctopp_c_scale), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(title = "CTOPP Color Scaled",
       x = "Grade",
       y = "Scaled Score (s)") +
  theme_classic()

b_cl2 <- ggplot(b_clean, aes(grade, ctopp_l_scale), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(title = "CTOPP Letter Scaled",
       x = "Grade",
       y = "Scaled Score (s)") +
  theme_classic()

box_c <- ggarrange(b_cc1, b_cl1, b_cc2, b_cl2,
                  ncol = 2,
                  nrow = 2,
                  legend = "none")
box_c
```

### CTOPP Scatterplots 

```{r ctopp scatterplots, echo = FALSE, message = FALSE, warning = FALSE}

# ctopp color plots
s_cc1 <- ggplot(b_clean, aes(age, ctopp_c_raw, color = grade)) +
  geom_point(alpha = 0.6) +
  labs(title = "CTOPP Color Raw Scores",
       x = "Age",
       y = "Seconds") +
  theme_classic()
 
s_cc2 <- ggplot(b_clean, aes(age, ctopp_c_scale, color = grade)) +
  geom_point(alpha = 0.6) +
  labs(title = "CTOPP Color Scaled Scores",
       x = "Age",
       y = "Scaled Score") +
  theme_classic()

# ctopp letter plots 
s_cl1 <- ggplot(b_clean, aes(age, ctopp_l_raw, color = grade)) +
  geom_point(alpha = 0.6) +
  labs(title = "CTOPP Letter Raw Scores",
       x = "Age",
       y = "Seconds") +
  theme_classic()

s_cl2 <- ggplot(b_clean, aes(age, ctopp_l_scale, color = grade)) +
  geom_point(alpha = 0.6) +
  labs(title = "CTOPP Letter Scaled Scores",
       x = "Age",
       y = "Scaled Score") +
  theme_classic()

# format figure
scatter_c <- ggarrange(s_cc1, s_cl1, s_cc2, s_cl2,
                  ncol = 2,
                  nrow = 2,
                  common.legend = TRUE)
scatter_c
```

### Color and letter raw scores overall sample and grade
# maybe add r on here

```{r}
s_cl_all <- ggplot(b_clean, aes(ctopp_c_raw, ctopp_l_raw)) +
  geom_point(alpha = 0.5) +
  labs(title = "CTOPP Raw Scores",
       x = "Color Naming (s)",
       y = "Letter Naming (s)") +
  theme_classic()

s_cl_all
```
# grade only plots
```{r}
b_clean %>% 
  filter(grade == "K") %>% 
  ggplot(aes(ctopp_c_raw, ctopp_l_raw)) +
  geom_point(alpha = 0.5) +
  labs(title = "Kindergarten CTOPP Raw Scores",
       x = "Color Naming (s)",
       y = "Letter Naming (s)") +
  theme_classic()

b_clean %>% 
  filter(grade == "1") %>% 
  ggplot(aes(ctopp_c_raw, ctopp_l_raw)) +
  geom_point(alpha = 0.5) +
  labs(title = "1st Grade CTOPP Raw Scores",
       x = "Color Naming (s)",
       y = "Letter Naming (s)") +
  theme_classic()

b_clean %>% 
  filter(grade == "2") %>% 
  ggplot(aes(ctopp_c_raw, ctopp_l_raw)) +
  geom_point(alpha = 0.5) +
  labs(title = "2nd Grade CTOPP Raw Scores",
       x = "Color Naming (s)",
       y = "Letter Naming (s)") +
  theme_classic()
```



## TOWRE

### TOWRE Histograms

```{r towre swe hist, echo = FALSE, message = FALSE, warning = FALSE}
# TOWRE_SWE 
h_ts1 <- ggplot(b_clean, aes(swe_raw)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 5, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(title = "TOWRE SWE",
       x = "# Correct",
       y = "Density") +
  theme_classic()


h_ts2 <- ggplot(b_clean, aes(swe_age)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 5, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) + 
  labs(x = "# Correct (Age Scaled)",
       y = "Density") +
  theme_classic()

h_ts3 <- ggplot(b_clean, aes(swe_grade)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 5, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(x = "# Correct (Grade Scaled)",
       y = "Density") +
  theme_classic()

```


```{r towre pde hist,  echo = FALSE, message = FALSE, warning = FALSE}

# TOWRE_PDE
h_tp1 <- ggplot(b_clean, aes(pde_raw)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 5, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(title = "TOWRE PDE",
       x = "# Correct",
       y = "Density") +
  theme_classic()

h_tp2 <- ggplot(b_clean, aes(pde_age)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 5, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(x = "# Correct (Age Scaled)",
       y = "Density") +
  theme_classic()

h_tp3 <- ggplot(b_clean, aes(pde_grade)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 5, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(x = "# Correct (Grade Scaled)",
       y = "Density") +
  theme_classic()

hist_t <- ggarrange(h_ts1, h_tp1, h_ts2, h_tp2, h_ts3, h_tp3,
                    ncol = 2,
                    nrow = 3,
                    common.legend = TRUE)
hist_t
```

### TOWRE Boxplots 

```{r towre swe boxplots, echo = FALSE, message = FALSE, warning = FALSE}
b_ts1 <- ggplot(b_clean, aes(grade, swe_raw), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(title = "TOWRE SWE",
       x = "Grade",
       y = "# Correct") +
  theme_classic()

b_ts2 <- ggplot(b_clean, aes(grade, swe_age), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(x = "Grade",
       y = "# Correct (Age Scaled)") +
  theme_classic()

b_ts3 <- ggplot(b_clean, aes(grade, swe_grade), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(x = "Grade",
       y = "# Correct (Grade Scaled)") +
  theme_classic()
```

```{r towre pde boxplots, echo = FALSE, message = FALSE, warning = FALSE}
b_tp1 <- ggplot(b_clean, aes(grade, pde_raw), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(title = "TOWRE PDE",
       x = "Grade",
       y = "# Correct") +
  theme_classic()

b_tp2 <- ggplot(b_clean, aes(grade, pde_age), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(x = "Grade",
       y = "# Correct (Age Scaled") +
  theme_classic()

b_tp3 <- ggplot(b_clean, aes(grade, pde_grade), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(x = "Grade",
       y = "# Correct (Grade Scaled)") +
  theme_classic()
```

```{r towre boxplot figure, echo = FALSE, fig.height = 6, warning = FALSE}
box_t <- ggarrange(b_ts1, b_tp1, b_ts2, b_tp2, b_ts3, b_tp3,
                  ncol = 2,
                  nrow = 3,
                  legend = "none")
box_t
```


### TOWRE Scatterplots
```{r towre swe scatterplots,  echo = FALSE, message = FALSE, warning = FALSE}
# SWE plots
t1 <- ggplot(b_clean, aes(age, swe_raw, color = grade)) +
  geom_point(alpha = 0.7) +
  labs(title = "TOWRE SWE Raw Scores",
       x = "Age",
       y = "# Correct") +
  theme_classic()

t3 <- ggplot(b_clean, aes(age, swe_age, color = grade)) +
  geom_point(alpha = 0.7) +
  labs(title = "TOWRE SWE Age-Scaled Scores",
       x = "Age",
       y = "# Correct (Age-Scaled)") +
  theme_classic()

t5 <- ggplot(b_clean, aes(age, swe_grade, color = grade)) +
  geom_point(alpha = 0.7) +
  labs(title = "TOWRE SWE Grade-Scaled Scores",
       x = "Age",
       y = "# Correct (Grade-Scaled)") +
  theme_classic()

```

```{r towre pde histograms,  echo = FALSE, message = FALSE, warning = FALSE}
t2 <- ggplot(b_clean, aes(age, pde_raw, color = grade)) +
  geom_point(alpha = 0.5) +
  labs(title = "TOWRE PDE Raw Scores",
       x = "Age",
       y = "# Correct") +
  theme_classic()

t4 <- ggplot(b_clean, aes(age, pde_age, color = grade)) +
  geom_point(alpha = 0.5) +
  labs(title = "TOWRE PDE Age-Scaled Scores",
       x = "Age",
       y = "# Correct (Age-Scaled)") +
  theme_classic()

t6 <-  ggplot(b_clean, aes(age, pde_grade, color = grade)) +
  geom_point(alpha = 0.5) +
  labs(title = "TOWRE PDE Grade-Scaled Scores",
       x = "Age",
       y = "# Correct (Grade-Scaled)") +
  theme_classic()
```

```{r towre scatterplot figure, echo = FALSE, fig.height = 8, warning = FALSE}
# format plots
scatter_towre <- ggarrange(t1, t2, t3, t4,t5, t6,
                  ncol = 2,
                  nrow = 3,
                  common.legend = TRUE)
scatter_towre
```

## Woodcock Johnson

### WJ Histograms

```{r wj histogram, echo = FALSE, warning = FALSE}
# WJ_NUM
h_wj1 <- ggplot(b_clean, aes(wj_raw)) +
  geom_histogram(aes( y = ..density..),
                 binwidth = 2, 
                 color = "black",
                 fill = "white") +
  geom_density(aes(fill = grade), alpha = 0.3) +
  labs(title = "Woodjcock Johnson Raw (basal rule)",
       x = "# Correct",
       y = "Density") +
  theme_classic()
```

### WJ Boxplot

```{r wj boxplot, echo = FALSE, warning = FALSE}
b_wj1 <- ggplot(b_clean, aes(grade, wj_raw), fill = grade) +
  geom_boxplot(aes(fill = grade),
               alpha = 0.5) +
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               dotsize = .5,
               fill = "white") +
  labs(x = "Grade",
       y = "# Correct") +
  theme_classic()
```

### WJ Scatterplot
```{r wj scatterplot, echo = FALSE, warning = FALSE}
# Still need: WJ age equivalent and WJ grade equivalent
s_wj1 <- ggplot(b_clean, aes(age, wj_raw, color = grade)) +
  geom_point(alpha = 0.7) +
  labs(x = "Age",
       y = "# Correct") +
  theme_classic()

```

```{r wj plots, fig.width = 10, echo = FALSE, warning = FALSE}
plots_wj <- ggarrange(h_wj1, b_wj1, s_wj1,
                      nrow = 1,
                      ncol = 3,
                      common.legend = TRUE)
plots_wj
```

## Full Correlation Matrix

```{r create data for corr matrix, echo = FALSE, warning = FALSE}
# select only numeric columns to use for correlation matrix (add WJ scaled scores later)
b_corr <- b_clean %>% 
  select(ctopp_c_raw,
         ctopp_c_scale,
         ctopp_l_raw,
         ctopp_l_scale,
         swe_raw,
         swe_age,
         swe_grade,
         pde_raw,
         pde_age,
         pde_grade,
         wj_raw,
         age) %>% 
  na.omit()
```

```{r correlation matrix, echo = FALSE, warning = FALSE}

# create correlation matrix (cor from package corrplot)
b_matrix <- cor(b_corr)

# something I found online to be able to see values in correlation matrix 
library(Hmisc)
res <- rcorr(as.matrix(b_corr))

# custom function I copied and pasted - need to find link
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

```

```{r correlogram, echo = FALSE, warning = FALSE}

# combine correlogram with significance test
b_test <- cor.mtest(b_corr, conf.level = .95)
corrplot(b_matrix, 
         p.mat = b_test$p, 
         insig = "label_sig",
         sig.level = c(.001, .01, .05),
         pch.cex = .7, 
         pch.col = "white",
         tl.col = "black")

# calling function from above to show correlation matrix values
corr_mat <- flattenCorrMatrix(res$r, res$P)
kable(corr_mat,
      digits = 3, # 3 digits (or however many you want)
      format = "latex",
      caption = "K2 Behavioral Descriptive Statistics") %>%
  kable_styling(latex_options = c("striped", "scale_down"),
                font_size = 4)
```
 
##Correlation Matrix (Raw Scores Only)

The histograms of the variables are shown on the diagonal. The asterisks indicate the significance levels of the correlations. Each significance level is associated to a symbol (from http://www.sthda.com/english/wiki/correlation-matrix-formatting-and-visualization/) :

p-values(0, 0.001, 0.01, 0.05, 0.1, 1) <=> symbols(“***”, “**”, “*”, “.”, " “)
 
```{r matrix hist, echo = FALSE, message = FALSE, warning = FALSE}
# something else I found online to plot correlations, histograms, and 
library("PerformanceAnalytics")
raw_perform <- b_corr %>% 
  select(ctopp_c_raw,
         ctopp_l_raw,
         swe_raw,
         pde_raw,
         wj_raw,
         age) %>% 
  mutate(ctopp_c_raw = ctopp_c_raw*(-1),
         ctopp_l_raw = ctopp_l_raw*(-1))

raw_matrix <- cor(raw_perform)
chart.Correlation(raw_matrix, histogram=TRUE, pch=19)
```
 
## Corrplot - Raw Scores Only

```{r}
raw_test <- cor.mtest(raw_perform, conf.level = .95)
corrplot(raw_matrix, 
         p.mat = raw_test$p, 
         insig = "label_sig",
         sig.level = c(.001, .01, .05),
         pch.cex = .7, 
         pch.col = "white",
         tl.col = "black")
```

 
# Basic Significance Tests - Raw Scores Only

*Tukey:* 
diff = difference between means of the two groups
lwr, upr = the lower and upper end point of the confidence interval
p adj = p-value after adjustment for multiple comparisons
 
```{r ctopp c aov}
ctopp_c_aov <- aov(ctopp_c_raw ~ grade, data = b_clean)
summary(ctopp_c_aov)
TukeyHSD(ctopp_c_aov) 
```

K and 2 are signficantly different from each other, but K is not significantly different from 1 and 1 is not significantly different from 2 at $\alpha$ = .05. 

```{r ctopp l aov}
ctopp_l_aov <- aov(ctopp_l_raw ~ grade, data = b_clean)
summary(ctopp_l_aov)
TukeyHSD(ctopp_l_aov)
```
K is sig. diff. from 1 and 2 but 1 and 2 are not sig diff from each other at $\alpha$ = .05.

```{r swe raw}
swe_aov <- aov(swe_raw ~ grade, data = b_clean)
summary(swe_aov)
TukeyHSD(swe_aov)
```
K is sig. diff. from 1 and 2 but 1 and 2 are not sig diff from each other at $\alpha$ = .05.

```{r pde raw}
pde_aov <- aov(pde_raw ~ grade, data = b_clean)
summary(pde_aov)
TukeyHSD(pde_aov)
```
All three grades are significantly different from each other at $\alpha$ = .05.

```{r wj raw}
wj_aov <- aov(wj_raw ~ grade, data = b_clean)
summary(wj_aov)
TukeyHSD(wj_aov)
```
All three grades are significantly different from each other at $\alpha$ = .05.

